---
title: "Chi-Square Test of Independence"
#author: "Burrel Vann Jr"
output: 
  html_document: 
    css: markdown3.css
---


###### Navigation

* [**Intro to R**](Intro-and-Univariate-Statistics.html)
* [**t-Test**](t-test.html) [[*t-Test example*]](t-test-example.html)
* [**Analysis of Variance (ANOVA)**](anova.html)
* [**Chi Square**](chi-square.html) 
* [**Correlation**](correlation.html) 


### What is the Chi Square Test of Independence?

The Chi Square test ($X^2$) examines the association or relationship between two nominal/ordinal variables to see if the relationship reflects a true relationship that we could expect to find in the population. The test also tells us whether or not a category (attribute) of one variable varies by categories of another variable. 

The test is called the *test of independence* because it really tests the absence of association between (independence of) two variables.





```{r, echo=F}
options(repos=c(CRAN="http://cran.stat.ucla.edu/"))
```

# 


```{r, results="hide", warning=FALSE, message=FALSE, echo=FALSE}
library(MASS)
library(psych)
library(lattice)
library(mosaic)
```

#

#### Assumptions and Diagnostics for Chi Square

The assumptions for the Chi Square are...

* Independence of Observations
* Normality

```{r, echo=FALSE}
data1 <- mtcars
```


##### 1. Independence of Observations (Examine Data Collection Strategy)
* Cases (observations) are not related or dependent upon each other. Case can’t have more than one attribute. No ties between observations. Examine data collection strategy to see if there are linkages between observations. 


##### 2. Normality (Examine Crosstabs for Expected Frequencies)


* Distributions must be relatively normal. A good way to see if we meet the normality assumption is to look at the **Expected Frequencies** in each cell of the crosstab. **If 20 percent or more of our Expected Frequencies across all the cells are less than 5 cases ($E \lt 5$), then you violate the assumption of normality**.

*The Chi Square test is finicky, in that you cannot see the expected frequencies (and observed frequencies) until you run the Chi Square function (see below)*


```{r, results="hide", echo=FALSE}
table(data1$vs, data1$am)
```

#
### The Chi Square Test Calculation

The calculation for the Chi Square is:

 $X^2 = \sum \frac{(f_o - f_e)^2}{f_e}$ or
 $X^2 = \sum \frac{(f_{o_i} - f_{e_i})^2}{f_{e_i}}$
 
where... <br>

* $f_o$ (or $f_{o_i}$) is the observed frequency for that cell (the $i^{th}$cell)<br>
* $f_e$ (or $f_{e_i}$) is the expected frequency for that cell (the $i^{th}$cell)<br>
  + $f_e$ (or $f_{e_i}$) for each cell is calculated by the following:
    + $f_e$ (or $f_{e_i}$) $= \frac{(r_{total_i})(c_{total_i})}{total}$ $= \frac{(n_{row_i})(n_{column_i})}{N}$


In addition, the degrees of freedom ($df$) for the test is...<br> 
* $df = (r-1)(c-1)$

where...

* $r$ is the number of rows in a crosstabulation<br>
* $c$ is the number of columns in a crosstabulation<br>



### Running the Chi Square Test



For Chi Square, within the <span style="color:blue">`chi.sq`</span> function, the dependent variable is listed first and the independent variable is listed second. 



```{r}
chisq.test(data1$vs, data1$am, correct=FALSE)
```

In the output above, we see the $X^2$-obtained value (0.90688), the degrees of freedom (30), and the p-value (0.3409, which is greater than our set alpha level of .05).

To interpret the findings, we report the following information:

* The test used
* If you **reject** or **fail to reject** the null hypothesis
* The variables used in the analysis
* The degrees of freedom, calculated value of the test ($X^2_{obtained}$), and $p-value$
  + $X^2(df) = X^2_{obtained}$, $p-value$

“Using the Chi Square test of independence ($X^2$), I reject/fail to reject the null hypothesis that there is no association between variable one and variable 2, in the population, $X^2(?) = ?, p ? .05$” 

“Using the Chi Square test of independence ($X^2$), I reject the null hypothesis that there is no association between the shape and the transmission type of cars, in the population, $X^2(1) = .90688, p \gt .05$” 


##### Back to the Normality Assumption

To check if you've met the assumption of normality (e.g. fewer than 20% of cells in the crosstab of expected frequencies falls below $n=5$), you have to make the chi square function (<span style="color:blue">`chi.sq`</span>) into an object... as such... 

```{r}
x2 <- chisq.test(data1$vs, data1$am, correct=FALSE)
```

and because the actual test is now an object, you can call it's component variables by using the object name, the dollar sign operator (<span style="color:blue">`$`</span>), and the variable named **expected** (you can also call the variable named **observed**) to get a crosstab of each. 

```{r}
x2$observed
```

```{r}
x2$expected
```

We see here that we have met the assumption of normality: fewer than 20% of cells in the 4x4 table have less than 5 expected counts (actually, no cell has fewer than 5).

<br><br><br>
